---
title: "p8105_hw5_rs4338"
author: "Rebecca Shyu"
date: "2024-11-12"
output: github_document
---

```{r setup, include=FALSE}
set.seed(1)

library(tidyverse)
library(broom)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%",
	dpi=300
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d



```


```{r prob1_function}
birthday = function(n) {
  if (!is.numeric(n)) {
    stop("Input (group size) needs to be numeric")
  }
  
  bday_vec = sort(round(runif(n, min = 1, max = 365)))
  
  duplicates = sum(duplicated(bday_vec)) > 0

  return(duplicates)
}
```

Based on the following plot looking at probability of at least two people in the group sharing a birthday as a function of group size, as the group size increases, the probability gets closer to 1. This makes sense because 0 to 50 is getting closer to the number of days in a year (365 in this case) and we're assuming that birthdays are uniformly distributed (not true in reality) so the probability increases as we draw more.

```{r prob1_sim}
bday_sim =
  expand_grid(
    n = 2:50,
    iter = 1:10000
  ) %>% 
  mutate(
    results = map_lgl(n, birthday)
  ) %>% 
  group_by(n) %>% 
  summarize(prob = mean(results))

bday_sim %>% 
  ggplot(aes(x = n, y = prob )) + 
  geom_line() +
  labs(
    title = "Probability of >1 Shared Bday Across Group Size",
    x = "Group Size (n)",
    y = "Probability"
    ) +
  theme(plot.title = element_text(hjust = 0.5))

```



```{r prob2_setup}
prob2_sim = function(mu) {
  
  n = 30
  alpha = 5

  sim_data = 
    tibble(
      x = rnorm(n, mean = mu, sd = alpha)
    )
  
  t_results = tidy(t.test(sim_data))
  
  out_df =
    tibble(
      p_val = t_results %>% pull(p.value),
      estimate_val = t_results %>% pull(estimate)
    )
  
  return(out_df)
}


prob2_results =
  expand_grid(
    mu = 0:6,
    iter = 1:5000
  ) %>% 
  mutate(
    t_tests = map(mu, prob2_sim)
  ) %>% 
  unnest(t_tests)
```

Based on the following plot, as mu (effect size) increased from 0 to 6, the power (proportion of iterations rejecting the null - p-value less than 0.05) also increased. The larger the effect size, the easier/larger likelihood for the t-test to reject the null hypothesis.

```{r prob2_effvspower}
prob2_results %>% 
  group_by(mu) %>% 
  summarise(proportion_reject = mean(p_val < 0.05)) %>% 
  ggplot(aes(x = mu, y = proportion_reject)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Effect Size vs Power",
    x = "Mu",
    y = "Proportion of Rejecting Null"
    ) +
  theme(plot.title = element_text(hjust = 0.5))
```


This plot shows the average estimates of mu for all samples (black) and only sames for which the null was rejected (red) vs the true value of mu (0-6). I've overlayed both onto one plot for easy comparison. The black dots are very close to the like y=x, but the red varies a lot until mu = 4 when it matches up with the black dots. This matches with the plot right above where at mu=4, the power is almost to 1 and is approximately equal to the true value of mu. Mu = 0 is interesting for the red dot because the averaged estimates also includes negative numbers which shows that it's closer to the actual value, but it has very low power in reality. 

```{r prob2_avgvstruemu}
avg_est_all = 
  prob2_results %>% 
  group_by(mu) %>% 
  summarise(avg_mu_est = mean(estimate_val))

avg_est_reject = 
  prob2_results %>% 
  filter(
    p_val <= 0.05
  ) %>% 
  group_by(mu) %>% 
  summarise(avg_mu_est = mean(estimate_val)) 

ggplot() +
  geom_point(data = avg_est_all, aes(x = mu, y = avg_mu_est, color = "All Estimates")) +
  geom_point(data = avg_est_reject, aes(x = mu, y = avg_mu_est, color = "Only Rejected Estimates")) +
  scale_color_manual(values = c("All Estimates" = "black", "Only Rejected Estimates" = "red")) +
  labs(
    title = "Average Estimates vs True Mus",
    x = "True Mu",
    y = "Average Mu Estimate",
    color = "Estimate Type"
    ) +
  theme(plot.title = element_text(hjust = 0.5))
```


